Sustainable Pathways for Irish homes in changing TIMES

### 2021-06-25 manuscript.tex
25 June 2021

In Ireland, previous modelling on the residential sector has often involved bottom-up residential models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland. In 2011 Dineen and Ó Gallachóir \cite{Dineen2011ModellingHeating} developed a residential bottom-up model to assess the impacts of measures in Ireland's National Energy Efficiency Action Plan (NEEAP). The model focused on energy efficiency of space and water heating by examining technology and building fabric efficiency. The building fabric efficiency of newly occupied dwellings developed takes into account the theoretical increase in energy efficiency due to improved 2008 and 2010 building regulations. Space heating demand accounted for 70\% of household energy consumption in newly built dwellings in 1997, down to just under 60\% in 2006 and the model is predicts a reduction to 27\% by 2015.The paper acknowledges retrofitting of existing stock will play a large role in meeting energy efficiency targets, however it is not examined. Nor is the affect of internal temperatures on energy efficiency improvements.  
\par In 2013 Ahern et al.\cite{Ahern2013StateStock} investigated thermal retrofit measures to the existing detached, oil centrally heated, rural housing stock. A base geometry according to age bands and thermal characteristics was created in the study. The heat load for a statistical occupancy and year was established. The default values used in some of the BER database variables where highlighted. For example,  there are assumed U-values when the roof insulation thickness is unknown, this is based on age, but there are no U-values for uninsulated roofs and a 2001 survey found 18\% of detached housing in Ireland had no roof insulation. The internal temperature assumed for all dwelling was 20$^{\circ}$C. Ahern et al. found that newer dwellings, consume the greatest amount of energy in kWh/annum than all the older dwellings, due to the fact that their large floor area. Concluding the majority of energy efficiency progress for thermal use in Irish dwellings is being offset.

The reduction in heating costs and CO2 emissions for houses constructed prior to 1979 (coinciding with the introduction of building regulations) and around 26% for newer homes, thus offering a significant contribution (44%) to Ireland’s residential carbon abatement projections and hence in meeting the EU’s directives on energy and carbon.

through energy efficiency measures has been offset by the so called size effect. The
The model predicts a reduction of 65% in running costs and
CO2 emissions for house types A–E and a reduction of 26% in running costs and CO2 emissions for house types F–J from retrofit measures, and forecasts a total saving of 276 million euro and a CO2 abatement of 1.21 MtCO2. By
the
potential exists to reduce running costs and CO2 emissions by an average of 63% for housing constructed prior to the 1979 building regulations (house types A–E) and by 26% for newer housing. The ratio of amounts of CO2 which can be potentially saved to
the amount of money saved was calculated as 4.35 kgCO2/h. Nationally an annual saving of h183 million euro is theoretically achievable with a resultant saving of 0.80 MtCO2 arising from insulation retrofits, the replacement of single glazed windows and increasing air tightness. The
\cite{Ahern2015EnergySystems}\cite{Dineen2015ImprovedSource}\cite{Uidhir2020ResidentialIreland}\cite{Uidhir2020ImprovingDecisions} and some models have been used to optimise Ireland's entire energy system \cite{Chiodi2013ModellingSystem}\cite{Connolly2021AalborgEnergy-System}, but very few models have used energy system optimisation model to explore the Irish residential sector. In 2014, Rogan et al used OSeMOYOS to explore energy service demands \cite{Rogan2014LEAPsSystem}, and Thellufsen examining DH with EnergyPLAN and TIMES in 2019 \cite{Thellufsen2019ImplementingIreland}.

@@

The Archetype Dwelling Energy Model SQL (ArDEM-SQL) is a simulation model built on IS EN 13790, it converts the BER public database into Microsoft SQL language and then calculates annual energy consumption for space heating, water heating, pumps \& fans and lighting. ArDEM-SQL is used to clean the input data in TIM, by changing internal temperatures. The internal temperature is highly dependent on occupancy behaviour, a thermal comfort  thermal discomfort are higher among the over-65s than for other household \cite{Healy2002FuelIreland} - living room temperatues between 12 and 26...s  and therefore there is are large variations per BER rating.
The International Standard ISO 7730 [6] uses these PMV and PPD indices to predict the thermal sensation of people exposed to moderate thermal environments, as well as to specify acceptable thermal environmental conditions for comfort...
Despite this, it is
worth noting that the most inefficient dwellings i.e. band F, managed to achieve an
average temperature during period 2 that was only 0.5°C lower than the most
efficient dwellings i.e. band C.
C = 19.2, D = 18, E = 18.2 and F= 18.7 older occupants dublin Ho
%Healy & Clinch reported that fuel-poor households were more likely to be experiencing colder temperatures than other households. In total almost 30% of fuelpoor households had living room temperatures below 18°C and more than two thirds had living room temperatures below 20°C. Overall nearly half of all non-fuel-poor homes had living room temperatures below 20°C. In households with persons aged over 65 years, 16% had living room temperatures below 18°C and over 50% had temperatures below 20°C. On the other end of the scale over 5% of households occupied by persons over 65 years experienced living room temperatures of ≥24°C.https://arrow.tudublin.ie/cgi/viewcontent.cgi?article=1089&context=scienmas
%The BER for the sample dwellings ranged from C1 to F. Over two thirds of the
sample dwellings had a BER of E or F. The remaining dwellings had a BER of C or
D. In total 4 of the sample dwellings had undergone energy efficiency upgrades in
recent years, including cavity wall insulation, attic insulation where applicable and
heating control upgrades. In the case of two sample dwellings, these upgrade
works were completed between monitoring period 1 and 2, and therefore results
could be compared before and after retrofit.

@@

Theoretically G-rated BERs consume more energy and consequently have higher energy bills to maintain a comfortable internal temperature, but in reality G-rated dwellings have low internal temperatures, which is a reason for Ireland's high excessive winter mortality \cite{Clinch2000HousingMortality}. Fuel Poverty which which is defined as when a \textit{``household spends 10\% or more of income to achieve adequate warmth"} is a major contributor to the high excessive winter deaths. An Element Energy report claimed 28\% of Irish households were in fuel poverty in 2015  \cite{ElementEnergy2015Bottom-upIreland}. Fuel Allowance and the Household Benefits provides support to low income households to alleviate fuel poverty. These payments have been increased over time with a further increase made in Budget 2020, coinciding with the increase in the carbon tax \cite{,EuropeanUnion2020EUObservatory}. Kerr's study finds an explicit link between fuel poverty and social welfare policy in Ireland \cite{,EuropeanUnion2020EUObservatory,Kerr2019PoliticsFrance} and O’Meara highlights the health issues associated with Fuel Poverty \cite{OMeara2016AIreland}. Ironically, it can be argued this places low income households deeply dependent on fossil fuels for health.
\par BERs do not take account of actual usage based on occupancy behaviour or occupancy density, nor does it take account of the size of the dwelling. The BER labels do therefore leave a gap between actual consumption and expected consumption, the literature review explains how this problem was tackled.
\subsubsection{House Stock Data}
\label{LitR:StockData}
EPC register EU - Ireland
%The quality of Swedish EPC data was initially studied by Göransson [41], subsequently by Claesson [42], Mangold et al. [21] and Hårsman et al. [43], and most recently by Pasichnyi et al. [6]. These studies have found a couple of important deficiencies in Swedish EPC data. First, the energy expert, and the firm he or she belongs to, has been shown to generate variations in Swedish EPC data [43], [44]. More specifically, Hårsman et al. [43] found that the energy expert along with the characteristics of the firm caused a variation of ±20% in the assessment of energy consumption.
%Second, through interviews with energy experts, Claesson [42] found that most EPCs contain uncertain estimates of energy use as well as arbitrarily distributed values—thus challenging the notion that Swedish EPC data quality, while allegedly measured, should be superior to the quality of calculated values in other member states.
CIARAs Paper
Every BER displays over 200 energy related variables per dwelling, and over half of all residential dwellings are accounted. This large and detailed BER database does
BER Database, CSO and Calculations and Filtration 
FIGURE - Expected and Actual Consumption per m2
Than Actual 
DEAP - The space heat use is defined in IS EN 13790 as the heat delivered to the heated space by an ideal heating
system to maintain the set-point temperature during a given period of time. The DEAP calculation of space
heat use is done on a monthly basis based on the procedure described in IS EN 13790. A single-zone
calculation is used, using a single average value of mean internal temperature as described in Section 7.In DEAP, the heating season is defined as running from October to May inclusive. The annual space heat use
is the sum of monthly values for these eight months.
The heating hours and required internal temperatures in DEAP are based on the requirements of a typical
household. The schedule is as follows.
• Weekdays: 07.00 to 09.00 and 17.00 to 23.00
• Weekends: 07.00 to 09.00 and 17.00 to 23.00
This standardised schedule represents a total heating period of 56 hours per week.
\subsubsection{Costs}
\label{LitR:Costs}
Elephant in the room - Fuel Poverty
%JF -The tax / subsidy inclusive prices are important in driving behaviour, and taxes and subsidies will be a vital policy instrument. Thus, uptake of technologies such as retrofitting will be driven, inter alia, by the net price for consumers.
\subsubsection{Savings}
\label{LitR:Savings}
\subsubsection{Energy Service Demands Projections}
\label{LitR:Demands}
Macroeconomic 
no elasticity - fuel prices and weather compensation?
Time slices - 40 etc. 
% textit{To accommodate an expanding population, the number of dwellings has also increased to currently stand at 1.7 million households, although this fgure remains signifcantly short of projected housing need, withannual average demand estimated to be up to 36,000 units for the next 30 years (IBEC 2018). Yet despite the number of new homes built to ever increasing energy performance standards, the Irish housing stock is among the poorest in Europe in terms of energy effciency (Goggins et al. 2016). Current trends also show that households are getting bigger, with an average 15\% increase in foor area across all homes between 2000 and 2016 (SEAI 2018) \cite{FahyEnergyPractice}}
%
% The growth rate of the housing stock in the period to 2030 is obtained from the ESRI's macroeconomic estimates \citep{Yakut2020}. 
% A multiple linear regression is calculated to project future housing stock based on population size and occupancy rate beyond 2030. The total housing stock is expected to increase by 79\% in 2050 relative to 2018 to accommodate the growing population and declining occupancy rates. The share of apartments changes from 12\% in 2018 to 24\% in 2050, while the shares of attached and detached dwellings decline by 4\% and 8\% respectively.
Historical population estimates and future projections are obtained from the Central Statistics Office (CSO), Ireland \citep{CentralStatisticsOffice2020}. We use the M2F1 scenario since it represents medium growth in population and is in line with population projections used in other national sources \citep{Yakut2020}.

@@

\subsection{Limitations}
\label{Dis:Limit}
Consumers’ final investment decision can be
represented by “willingness-to-pay” curves – for
instance, around 10\% of owner-occupiers are willing
to invest in energy efficiency measures for a payback
time of 4 years, falling to 0\% for a payback time of
6 years for some consumer cohorts
Type of Tenure is ignored
BER aggeragateed
housing stock u-values
behaviour and occupancy
Area
\section{Conclusions}
\label{Conclusion}
This paper reports results on ambitious mitigation target in the period to 2050 for the Irish energy system. The analysis have been performed using the Irish TIMES model
\textit{We stress the importance of differentiating between technical con-
straints and behavioural constraints in optimisation models. Including both conflates what is considered desirable under the economic welfare paradigm (i.e. cost-minimising decisions) with “realistic decisions”, complicating the interpretation of the results. Optimisation models are by design well suited to study cost-optimal behaviour, and running scenarios just with technical constraints facilitates the identification of areas where market barriers are more entrenched, and interventions to accelerate the uptake of low-carbon practices may be needed.} \cite{Astudillo2017CanModelb}
\section{Acknowledgements}
\label{sec:Acknowledgements}
Tomas %CAPACITY is funded under the Climate Action Modelling Group (CAMG) by DECC
% https://www.marei.ie/project/capacity/
CHIMERA is supported by a research grant from Science Foundation Ireland (SFI) and the National Natural Science Foundation of China (NSFC) under the SFI-NSFC Partnership Programme, grant no. 17/NSFC/5181 and
supported by MaREI, the SFI Research Centre for Energy, Climate, and Marine [Grant No: 12/RC/2302\_P2]

>>>>

In Ireland, previous modelling on the residential sector has often involved bottom-up residential models, rather than energy system models specifically examining the residential sector. Nevertheless, the residential models have provided insights on input data and methodology approach used in Ireland. 
Early work by Healy and CLinch ---
2009 Hull

In 2011 Dineen and Ó Gallachóir \cite{Dineen2011ModellingHeating} developed a residential bottom-up model to assess the impacts of measures in Ireland's National Energy Efficiency Action Plan (NEEAP). The model focused on energy efficiency of space and water heating by examining technology and building fabric efficiency. The building fabric efficiency of newly occupied dwellings developed takes into account the theoretical increase in energy efficiency due to improved 2008 and 2010 building regulations. Space heating demand accounted for 70\% of household energy consumption in newly built dwellings in 1997, down to just under 60\% in 2006 and the model is predicts a reduction to 27\% by 2015.The paper acknowledges retrofitting of existing stock will play a large role in meeting energy efficiency targets, however it is not examined. Nor is the affect of internal temperatures on energy efficiency improvements.  
\par In 2013 Ahern et al.\cite{Ahern2013StateStock} investigated thermal retrofit measures to the existing detached, oil centrally heated, rural housing stock. A base geometry according to age bands and thermal characteristics was created in the study. The heat load for a statistical occupancy and year was established. The default values used in some of the BER database variables where highlighted. For example,  there are assumed U-values when the roof insulation thickness is unknown, this is based on age, but there are no U-values for uninsulated roofs despite a 2001 survey found 18\% of detached housing in Ireland had no roof insulation. Ahern et al. found dwellings built post-2000 consume the greatest amount of energy (kWh/annum), due to their large floor area. Concluding that the majority of energy efficiency progress for thermal use in Irish dwellings is being offset. The model thermal improvement measures resulted a reduction of running costs and CO2 emissions by an average of 65\% for houses constructed prior to 1979, this resulted in an average 12-year payback. Houses built in 1979 or after resulted in an average 26\% reduction in running costs and $CO_2$ emissions from retrofit measures.

\par In 2015 Dineen et al.\cite{Dineen2015ImprovedSource} built on previous work in examing new dwelling, to examining energy savings potential of energy efficiency improvement measures of the existing dwelling stock. The model relied on the BER database, which accounted for 18\% of all Irish dwellings, currently the database has more than doubled and accounts for over 50\%. The model used 175 archetype dwellings ( 5 building types, 7 energy bands and 5 wall construction types ) and 6 different retrofit measures were applied. BER database filters were used in this model, the BER filters from this study, were updated and applied in TIM. A sensitivity analysis is performed on the effects of internal temperature and rebound. It was found that there was approximately a 7.5\% change in energy consumption for every 1$^{\circ}$C internal temperature shift. Concluding, BER standard assumptions overestimate dwelling energy consumption and that the average internal temperature across the Irish dwelling stock may be closer to 18$^{\circ}$C than 21$^{\circ}$C.

Mac Uidhir et al.\cite{Uidhir2020ImprovingDecisions}  created Archetype Dwelling Energy Model - Structured Query Language (ArDEM-SQL), to simulate the potential for improved energy efficiency gains within the existing retrofit programs, specifically the Better Energy Homes (BEH) scheme.
BER database had grown to in ex- cess of 735,0 0 0 records
We introduce a new methodology that provides insights into sub-optimal retrofit choices.
The five most common retrofit combinations are simulated for each of the nine archetypes. The 3 energy rating types and 3 archetypes
The results show that the alternative retrofit combination differs by archetype and that additional energy effi- ciency gains of up to 86% can be achieved due to alternative retrofit choices
while maintaining the total number of retrofits completed, we estimate an energy savings potential of 1389 GWh, realising an additional 643 GWh or an 86% additional energy savings potential relative to those in the baseline data, see Fig. 7
The simulation of alternative archetype retrofit measures
yielded additional energy savings of 86% when compared with the observed energy savings as they appear in the baseline sce- nario. The additional energy savings are not evenly distributed across all building archetypes with detached CD/EFG and terraced CD/EFG account for 98% of the potential additional energy sav- ings when using the alternative retrofits identified, suggesting that greater energy savings are possible when retrofit measures are ap- plied to less energy efficient dwellings. It
\citep{Ahern2020ADatabases}
and some models have been used to optimise Ireland's entire energy system \cite{Chiodi2013ModellingSystem}\cite{Connolly2021AalborgEnergy-System}, but very few models have used energy system optimisation model to explore the Irish residential sector. In 2014, Rogan et al used OSeMOYOS to explore energy service demands \cite{Rogan2014LEAPsSystem}, and Thellufsen examining DH with EnergyPLAN and TIMES in 2019 \cite{Thellufsen2019ImplementingIreland}.

@@

The ArDEM-SQL is a simulation model built on IS EN 13790, it converts the BER public database into Microsoft SQL language and then calculates annual energy consumption for space heating, water heating, pumps \& fans and lighting. ArDEM-SQL is used to clean the input data in TIM, by changing internal temperatures. The internal temperature is highly dependent on occupancy behaviour, a thermal comfort  thermal discomfort are higher among the over-65s than for other household \cite{Healy2002FuelIreland} - living room temperatues between 12 and 26...s  and therefore there is are large variations per BER rating.
The International Standard ISO 7730 [6] uses these PMV and PPD indices to predict the thermal sensation of people exposed to moderate thermal environments, as well as to specify acceptable thermal environmental conditions for comfort...
Despite this, it is
worth noting that the most inefficient dwellings i.e. band F, managed to achieve an
average temperature during period 2 that was only 0.5°C lower than the most
efficient dwellings i.e. band C.
C = 19.2, D = 18, E = 18.2 and F= 18.7 older occupants dublin Ho
%Healy & Clinch reported that fuel-poor households were more likely to be experiencing colder temperatures than other households. In total almost 30% of fuelpoor households had living room temperatures below 18°C and more than two thirds had living room temperatures below 20°C. Overall nearly half of all non-fuel-poor homes had living room temperatures below 20°C. In households with persons aged over 65 years, 16% had living room temperatures below 18°C and over 50% had temperatures below 20°C. On the other end of the scale over 5% of households occupied by persons over 65 years experienced living room temperatures of ≥24°C.https://arrow.tudublin.ie/cgi/viewcontent.cgi?article=1089&context=scienmas
%The BER for the sample dwellings ranged from C1 to F. Over two thirds of the
sample dwellings had a BER of E or F. The remaining dwellings had a BER of C or
D. In total 4 of the sample dwellings had undergone energy efficiency upgrades in
recent years, including cavity wall insulation, attic insulation where applicable and
heating control upgrades. In the case of two sample dwellings, these upgrade
works were completed between monitoring period 1 and 2, and therefore results
could be compared before and after retrofit.

@@

Theoretically G-rated BERs consume more energy and consequently have higher energy bills to maintain a comfortable internal temperature, but in reality G-rated dwellings have low internal temperatures, which is a reason for Ireland's high excessive winter mortality \cite{Clinch2000HousingMortality}. Fuel Poverty which which is defined as when a \textit{``household spends 10\% or more of income to achieve adequate warmth"} is a major contributor to the high excessive winter deaths. An Element Energy report claimed 28\% of Irish households were in fuel poverty in 2015  \cite{ElementEnergy2015Bottom-upIreland}. Fuel Allowance and the Household Benefits provides support to low income households to alleviate fuel poverty. These payments have been increased over time with a further increase made in Budget 2020, coinciding with the increase in the carbon tax \cite{,EuropeanUnion2020EUObservatory}. Kerr's study finds an explicit link between fuel poverty and social welfare policy in Ireland \cite{,EuropeanUnion2020EUObservatory,Kerr2019PoliticsFrance} and O’Meara highlights the health issues associated with Fuel Poverty \cite{OMeara2016AIreland}. Ironically, it can be argued this places low income households deeply dependent on fossil fuels for health.
\par BERs do not take account of actual usage based on occupancy behaviour or occupancy density, nor does it take account of the size of the dwelling. The BER labels do therefore leave a gap between actual consumption and expected consumption, the literature review explains how this problem was tackled.
\subsubsection{House Stock Data}
\label{LitR:StockData}
EPC register EU - Ireland
%The quality of Swedish EPC data was initially studied by Göransson [41], subsequently by Claesson [42], Mangold et al. [21] and Hårsman et al. [43], and most recently by Pasichnyi et al. [6]. These studies have found a couple of important deficiencies in Swedish EPC data. First, the energy expert, and the firm he or she belongs to, has been shown to generate variations in Swedish EPC data [43], [44]. More specifically, Hårsman et al. [43] found that the energy expert along with the characteristics of the firm caused a variation of ±20% in the assessment of energy consumption.
%Second, through interviews with energy experts, Claesson [42] found that most EPCs contain uncertain estimates of energy use as well as arbitrarily distributed values—thus challenging the notion that Swedish EPC data quality, while allegedly measured, should be superior to the quality of calculated values in other member states.
CIARAs Paper
Every BER displays over 200 energy related variables per dwelling, and over half of all residential dwellings are accounted. This large and detailed BER database does
BER Database, CSO and Calculations and Filtration 
FIGURE - Expected and Actual Consumption per m2
Than Actual 
DEAP - The space heat use is defined in IS EN 13790 as the heat delivered to the heated space by an ideal heating
system to maintain the set-point temperature during a given period of time. The DEAP calculation of space
heat use is done on a monthly basis based on the procedure described in IS EN 13790. A single-zone
calculation is used, using a single average value of mean internal temperature as described in Section 7.In DEAP, the heating season is defined as running from October to May inclusive. The annual space heat use
is the sum of monthly values for these eight months.
The heating hours and required internal temperatures in DEAP are based on the requirements of a typical
household. The schedule is as follows.
• Weekdays: 07.00 to 09.00 and 17.00 to 23.00
• Weekends: 07.00 to 09.00 and 17.00 to 23.00
This standardised schedule represents a total heating period of 56 hours per week.

Scaling BER to CSO
\begin{equation} 
\begin{split}
S_y =  \frac{C_y} {B_y}
\end{split}
%\[ \sum_{n=1}^{\infty} 2^{-n} = 1 \]
\label{Eq:BERtoCSO}
\end{equation}
The number of dwellings by year of construction within the
BER database (B y ) is then scaled to align with the total number of dwellings as they appear in the CSO dataset (C y ) –for each archetype. Eq. (8) describes the scaling factor (S y ), used for each year of construction bracket

\begin{equation} 
\begin{split}
A_y =  {S_y} \cdot {B_{yarc}}
\end{split}
\label{Eq:BERtoCSO2}
\end{equation}
Number of archetypes dwellings by year of construction , where A k = Frequency of each archetype, S y = Scaling Factor, B yarc = Number Dwellings by BER category, A k = Frequency of each archetype.

Model calibration: archetype demand The
\subsubsection{Costs}
\label{LitR:Costs}
Elephant in the room - Fuel Poverty
%JF -The tax / subsidy inclusive prices are important in driving behaviour, and taxes and subsidies will be a vital policy instrument. Thus, uptake of technologies such as retrofitting will be driven, inter alia, by the net price for consumers.
\subsubsection{Savings}
\label{LitR:Savings}
\subsubsection{Energy Service Demands Projections}
\begin{equation} 
\begin{split}
E_y = {\sum_{A,M}} N_{A,y} \cdot S_{A,M}
\end{split}
%\[ \sum_{n=1}^{\infty} 2^{-n} = 1 \]
\label{Eq:RetrofitSaving}
\end{equation}
$E_y$ = Energy Savings in Year y
$N_{A,y}$ = Number of dwellings of archetype A retrofitted in year y
$S_{A,M}$ = ¼ Energy Savings per annum for retrofit measureM carried
out on archetype dwelling A
A = For each archetype
M = For each retrofit measure carried out

Next - 
\begin{equation} 
\begin{split}
D_M = {\sum_{A,M}^n} (A_kI_{ki} - A_kI_{kT} )
\end{split}
%\[ \sum_{n=1}^{\infty} 2^{-n} = 1 \]
\label{Eq:EnergyandRetrofitSaving}
\end{equation}
Total Energy Demand , where D T = Total Energy Demand, n = total number of archetypes, A k = Frequency of each archetype, I k = Energy Intensity of each archetype - D M = Energy Savings by retrofit measure (M), n = total number of archetypes, A k = Frequency of each archetype dwellings eligible for retrofit measure, I ki = Pre-retrofit archetype energy intensity, I kr - post-retrofit archetype intensity - Tomas Paper

Heterogeneous Occupancy Behaviour > House Stock 
\label{LitR:Demands}
Residential Energy was >50\%higher before internal temperature adjustments
Macroeconomic 
no elasticity - fuel prices and weather compensation?
Time slices - 40 etc. 
% textit{To accommodate an expanding population, the number of dwellings has also increased to currently stand at 1.7 million households, although this fgure remains signifcantly short of projected housing need, withannual average demand estimated to be up to 36,000 units for the next 30 years (IBEC 2018). Yet despite the number of new homes built to ever increasing energy performance standards, the Irish housing stock is among the poorest in Europe in terms of energy effciency (Goggins et al. 2016). Current trends also show that households are getting bigger, with an average 15\% increase in foor area across all homes between 2000 and 2016 (SEAI 2018) \cite{FahyEnergyPractice}}
%
% The growth rate of the housing stock in the period to 2030 is obtained from the ESRI's macroeconomic estimates \citep{Yakut2020}. 
% A multiple linear regression is calculated to project future housing stock based on population size and occupancy rate beyond 2030. The total housing stock is expected to increase by 79\% in 2050 relative to 2018 to accommodate the growing population and declining occupancy rates. The share of apartments changes from 12\% in 2018 to 24\% in 2050, while the shares of attached and detached dwellings decline by 4\% and 8\% respectively.
Historical population estimates and future projections are obtained from the Central Statistics Office (CSO), Ireland \citep{CentralStatisticsOffice2020}. We use the M2F1 scenario since it represents medium growth in population and is in line with population projections used in other national sources \citep{Yakut2020}.

@@

\subsection{Limitations}
\label{Dis:Limit}
Consumers’ final investment decision can be
represented by “willingness-to-pay” curves – for
instance, around 10\% of owner-occupiers are willing
to invest in energy efficiency measures for a payback
time of 4 years, falling to 0\% for a payback time of
6 years for some consumer cohorts
Type of Tenure is ignored
BER aggeragateed
housing stock u-values
behaviour and occupancy
Area
expected energy consumption weakness of BER database as 
Actual energy consumption for dwellings of
similar or near identical construction may vary by up to 600%
depending on usage \cite{Sunikka-Blank2012BuildingConsumption}
\section{Conclusions}
\label{Conclusion}
This paper reports results on ambitious mitigation target in the period to 2050 for the Irish energy system. The analysis have been performed using the Irish TIMES model
\textit{We stress the importance of differentiating between technical con-
straints and behavioural constraints in optimisation models. Including both conflates what is considered desirable under the economic welfare paradigm (i.e. cost-minimising decisions) with “realistic decisions”, complicating the interpretation of the results. Optimisation models are by design well suited to study cost-optimal behaviour, and running scenarios just with technical constraints facilitates the identification of areas where market barriers are more entrenched, and interventions to accelerate the uptake of low-carbon practices may be needed.} \cite{Astudillo2017CanModelb}
\section{Acknowledgements}
\label{sec:Acknowledgements}
Tomas %CAPACITY is funded under the Climate Action Modelling Group (CAMG) by DECC
% https://www.marei.ie/project/capacity/
CHIMERA is supported by a research grant from Science Foundation Ireland (SFI) and the National Natural Science Foundation of China (NSFC) under the SFI-NSFC Partnership Programme, grant no. 17/NSFC/5181 and
supported by MaREI, the SFI Research Centre for Energy, Climate, and Marine [Grant No: 12/RC/2302\_P2]
